Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2023-05-09 manuscript.Rmd
Updates

```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the average response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4). The grey lines represent the average within individual effects, and the blue line the aggregated effect over all individuals. Panel B shows how individual estimates are shrunken towards the general trend. Sample represents the observed effect per participant in seconds, while Model represents the estimated effects per participant. ", fig.height=3, fig.width=6, out.width ="95%"}
knitr::include_graphics("Images/Fig1_meanrtandhierplot_v2.png")
```

@@

If estimated for each participant independently, observed individual effects are quite variable because they are perturbed by sample noise. In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend [@efron1977stein]. This phenomenon, also called `shrinkage', is illustrated in Panel B of Figure \@ref(fig:figure1). The rightmost observation in the figure is in sample quite large compared to the other observed effects. The estimated effect (green point) is drastically corrected towards the mean. Conversely, the general trend is only influenced slightly by individuals showing a highly divergent trend. By adding the individual variability to the model, we no longer assume independence between observations, but the overall effect in the model leaves the possibility to generalize the effect. Therefore, hierarchical analysis can be considered a middle ground between an independent analysis and an analysis of aggregated data. 
```{r figure2, echo=FALSE, eval = FALSE, fig.cap="Sample represents the observed response time in milliseconds per individual, while model represents the individual response time estimates. Individual estimates are shrunken towards the general trend. ", fig.align = "left", fig.height = 2, fig.width = 3, out.width="70%"}
knitr::include_graphics("Images/hierplotefmor.png")
```

@@

```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and log-normal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
```

@@

```{r distritypeplot, echo=FALSE, fig.cap="Common probability density distributions that may be used as priors. Panel A shows normal distributions varying in mean and standard deviation. Panel B shows inverse gamma distributions that differ in the shape and scale parameter. Panel C shows Student's \\textit{t}-distributions that differ in the location and scale parameter. For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
knitr::include_graphics("Images/Fig4_distrtypefigure_v3.png")
```

@@

```{r priorpredfigurepanel, echo=FALSE, fig.cap="Side and digit effects from predicted data. Violins represent the distribution of predicted effects across 1000 simulation runs. The blue points are the average effects across simultions. Panel A shows the prior predictions for the normal model based on the default priors of \\textit{brms}. Panel B shows the prior predictions for the normal model based on the selected priors. Panel C shows the prior predictions for the log-normal model based on the priors chosen for the normal model. Panel D shows the prior predictions for the log-normal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
knitr::include_graphics("Images/Fig5_priorpredfigure_v2.png")
```

@@

The normal distribution on $\mu_{\beta}$ and $\mu_{\delta}$ seems reasonable if we assume that a two-sided hypothesis about the symbolic distance effect is tested. However, we expect less variance around this mean. Therefore, we decrease the setting for the variation. The wide tails of the Student's *t*-distribution could be one of the reasons for the high range in variation. We change the prior for $\mu_{\gamma}$ to a normal distribution that has more narrow tails. As RTs cannot be negative, we adjust the prior to be positive only by truncating the normal distribution using a lower bound of zero. Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution. ^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997]. For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.]\Revision{However, in `rstan`, non-conjugate priors are commonly applied [e.g., @gelman2006prior]} Therefore, we choose an inverse gamma distribution. We can also be more specific about the expected RTs. Based on our knowledge of these types of tasks [e.g., @haaf2017developing], it seems highly plausible that participants will respond within a second or two. We would expect effects in the range of 10 milliseconds to 100 milliseconds. We adjust our priors using these new distribution types and settings, evaluate the prior predictions (shown in panel B of Figure \@ref(fig:priorpredfigurepanel)) and arrive at the following priors:^[Because response times cannot be negative the prior for the overall baseline $\mu_\gamma$ is a truncated normal with a lower bound of zero.]

@@

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. \Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model [@fink1997compendium]. For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero. Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.} 
As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel). These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

```{r priorpredfigurepanellog, echo=FALSE, fig.cap="Predicted mean RT difference between digits per repetition. The blue line represents the aggregated trend. Panel A represents the prior prediction based on the priors of the normal model. Panel B shows the prior prediction based on the selected prior for the log-normal model.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%", eval = FALSE}
knitr::include_graphics("Images/priorpredfigure_lognormal_v2.png")
```
The transformation of the prior distributions is not as simple as the transformation of the parameter itself. Panel C shows that with the same priors as for the normal model, we obtain different data for the log-normal model compared to the normal model (panel B). Therefore, we adjusted the priors until the simulated data matched our expectations and looked like panel B. This yields the following priors that result in the prior prediction shown in panel D of Figure \@ref(fig:priorpredfigurepanel):^[For the overall baseline $\mu_\gamma$ we adjusted the prior based on our expectation of the expected overall response times. A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.]

@@

```{r traceplots3p, echo=FALSE, fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%", fig.cap="Panel A and B show the trace plot of the digit parameter $\\mu_{\\delta_{6}}$ for both packages. Panel C shows the estimated $\\hat{R}$ for all individual effects per parameter group. The stars represent the $\\hat{R}$ of the general effects. Panel D illustrates the variation in the estimated number of effective samples per parameter group. The dashed lines represent the total number of iterations for each package. The stars represent the number of effective samples of the general effects."}
knitr::include_graphics("Images/Fig8_traceplots3pnew_v2.png")
```

### Model Fitting 
The model is fitted using the packages `rstan` and `brms`. \Revision{We used different settings for `rstan` and `brms` due to the difference between the models for these packages. Although we tried to make the models as similar as possible, the models are not identical. This is partly due to difference in prior specification (i.e., specifying prior on the variances versus standard deviations).} For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 12000 samples from the joint posterior distribution. For `brms`, four chains are run with 6000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 20000 draws from the joint posterior distribution.

@@

```{r modelposteriorsmufigure3, echo=FALSE, fig.cap="Panel A shows the empirical mean RTs per digit. The standard error is represented by bars around the dots. Panel B and C show the posterior distributions for the general effects estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
knitr::include_graphics("Images/Fig9_posteriormu_v3.png")
```

@@

```{r randomparfigure, echo=FALSE, fig.cap="The posterior means for individual effect parameters with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
knitr::include_graphics("Images/Fig10_randompar_v2.png")

@@

```{r modelestimatesfigure3, echo=FALSE, fig.cap="Model estimates for digit effect parameters in \\textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates. ", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
knitr::include_graphics("Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
```

@@

```{r traceplotlog, echo=FALSE, fig.cap="(ref:captiontraceplotlog)", fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%"}
knitr::include_graphics("Images/Fig12_traceplots3plog_v2.png")
```

### Results
#### General effects  
We inspect the posterior distributions for the general effect parameters presented in Figure \@ref(fig:postplotlog). 
```{r postplotlog, echo=FALSE, fig.cap="The posterior distributions for the general effects in the log-normal model estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
knitr::include_graphics("Images/Fig13_posteriormulog_v2.png")
```

@@

```{r randomparplotlog, echo=FALSE, fig.cap="The posterior means for individual effect parameters in the log-normal model with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
knitr::include_graphics("Images/Fig14_randomparlog_v2.png")
```
We can further inspect the individual estimates for the digit effects in Figure \@ref(fig:modeldigitestplotlog). The figure shows the individual estimates for the digit effects with the distribution of the variability of the individual estimates (not the posterior distribution of the general effect). The figure illustrates that there is considerable variance in the point estimates of the individual digit effects. However, compared to the normal model there is considerably more hierarchical shrinkage. \Revision{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model. Or, the group mean in the log-normal model is not affected as much by individuals showing a divergent trend compared to the normal model. Whether we should prefer a model with more shrinkage depends on the theoretical considerations. If more individual variation is expected, less shrinkage would be preferred. However, the advantage of more shrinkage is that outliers do not influence the estimation of the effect as much. }

@@

```{r modeldigitestplotlog, echo=FALSE, fig.cap="Model estimates for digit effect parameters in the log-normal model in \\textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
knitr::include_graphics("Images/Fig15_inddiflogplot_onlyrstan_v2.png")
```

@@

The Bayes factor can also be interpreted as the relative evidence of two competing models. Bayes factors have a few very convenient characteristics. One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominatior in equation 13. For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al. [-@wagenmakers2018bayesian].

@@

Imagine this results in $\text{BF}_{\text{nl}} = 100$. This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model. \textcolor{blue}{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}. As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data. Therefore, Bayes factors are odds or ratios. Odds themselves are directly interpretable without the need for decisions or cutoffs. For example, if a presidential candidate is favored 10-to-1 over another one, then these are just the odds, and it is beside the point whether these are large or not. According to this line of thought, Rouder et al. [-@rouder2018theories] argue that researchers should take a similar approach when interpreting and reporting Bayes factors. While there exist rules of thumb to categorize Bayes factors verbally, we suggest researchers use their research context to form substantive conclusions based on Bayes factors, not arbitrary rules of thumb.

@@

```{r sdratioplotall, echo=FALSE, fig.cap="The prior and posterior distribution of the side parameter estimated by the \\textit{brms} package. The dots represent the height of the distributions at zero. The ratio of these dots indicates the Savage-Dickey density ratio. The Bayes factor represents the evidence for the hypothesis that the effect equals zero. A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.", fig.align = "left", fig.height = 6, fig.width = 4, out.width="60%"}
knitr::include_graphics("Images/Fig16_sdrplotside_v2.png")
```
With `rstan`, the SDD-ratio can be computed by fitting a density function to the posterior distribution obtained from the MCMC samples, and comparing the value of this density function at $\mu_{\beta} = 0$ to the value of the prior distribution at $\mu_{\beta} = 0$. This is illustrated in the code block below. First, we extract the posterior samples for the parameter $\mu_{\beta}$, the general effect of side. We apply the `logspline` function [@logspline] to the samples to obtain an estimate of the log-density of the posterior distribution. The density at the zero is calculated with the `dlogspline` function. Finally, we compare the estimated density of the posterior to the density of the prior. \textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} [-@van2021simulation].

@@

```{r bsbfestimatesbrmsrstan, echo=FALSE, fig.cap="The BF estimates of the other models (0; null, side and digit model) against the full model (1). The green triangles represent the estimates using \\textit{rstan}. The orange dots represent the \\textit{brms} estimates. The dashed line represents a $\\text{BF}_{01}$ of 1. ", fig.align = "left", fig.height = 11, fig.width = 4, out.width="70%"}
knitr::include_graphics("Images/Fig17_normalbfplot01.png")
```
Finally, it is time to answer our research question: Is there a symbolic distance effect? When considering all Bayes factor estimates from Figure \@ref(fig:sdratioplotall) and \@ref(fig:bsbfestimatesbrmsrstan), the null model ($\text{BF}_{max} = 232 * 10^{105}$ for `rstan`) performs best compared to the side model ($\text{BF}_{max} = 458 * 10^{68}$ for `rstan`), the digit model ($\text{BF}_{max} = 269 * 10^{31}$ for `rstan`). Therefore, the data from Rouder et al. (2005) show evidence against the symbolic distance effect.

@@

### Log-normal Model
We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity. We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model. This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences. For scenario 2, we remove individual variability from the side, digit, and full model. This resulted in the following posterior model probabilities shown in Table \@ref(tab:postmodproblog).

@@

The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability. This is in line with our findings for the normal model. Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood. The normal digit model from scenario 3 to the log-normal model from scenario 2. The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$. This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.

This example illustrates the benefits of comparing hierarchical models of varying complexity. While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary. As shown in Tables \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.

@@

This tutorial showed the application of two R packages: `rstan` and `brms`. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using `brms` because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and (virtually) always needs to be adapted. Additionally, users are limited to predefined parametrizations. `rstan` offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.

>>>>

```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the average response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4). The grey lines represent the average within individual effects, and the blue line the aggregated effect over all individuals. Panel B shows how individual estimates are shrunken towards the general trend. Sample represents the observed effect per participant in seconds, while Model represents the estimated effects per participant. ", fig.height=3, fig.width=6, out.width ="95%"}
knitr::include_graphics("I - Images/Fig1_meanrtandhierplot_v2.png")
```

@@

If estimated for each participant independently, observed individual effects are quite variable because they are perturbed by sample noise. In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend [@efron1977stein]. This phenomenon, also called `shrinkage', is illustrated in Panel B of Figure \@ref(fig:figure1). The rightmost observation in the figure is in sample quite large compared to the other observed effects. The estimated effect (green point) is drastically corrected towards the mean. Conversely, the general trend is only influenced slightly by individuals showing a highly divergent trend. By adding the individual variability to the model, we no longer assume independence between observations, but the overall effect in the model leaves the possibility to generalize the effect. Therefore, hierarchical analysis can be considered a middle ground between an independent analysis and an analysis of aggregated data. 
```{r figure2, echo=FALSE, eval = FALSE, fig.cap="Sample represents the observed response time in milliseconds per individual, while model represents the individual response time estimates. Individual estimates are shrunken towards the general trend. ", fig.align = "left", fig.height = 2, fig.width = 3, out.width="70%"}
knitr::include_graphics("I - Images/hierplotefmor.png")
```

@@

```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and log-normal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
knitr::include_graphics("I - Images/Fig2_Hierarchical structure.png")
```

@@

```{r distritypeplot, echo=FALSE, fig.cap="Common probability density distributions that may be used as priors. Panel A shows normal distributions varying in mean and standard deviation. Panel B shows inverse gamma distributions that differ in the shape and scale parameter. Panel C shows Student's \\textit{t}-distributions that differ in the location and scale parameter. For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
knitr::include_graphics("I - Images/Fig4_distrtypefigure_v3.png")
```

@@

```{r priorpredfigurepanel, echo=FALSE, fig.cap="Side and digit effects from predicted data. Violins represent the distribution of predicted effects across 1000 simulation runs. The blue points are the average effects across simultions. Panel A shows the prior predictions for the normal model based on the default priors of \\textit{brms}. Panel B shows the prior predictions for the normal model based on the selected priors. Panel C shows the prior predictions for the log-normal model based on the priors chosen for the normal model. Panel D shows the prior predictions for the log-normal model based on the selected priors.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%"}
knitr::include_graphics("I - Images/Fig5_priorpredfigure_v2.png")
```

@@

The normal distribution on $\mu_{\beta}$ and $\mu_{\delta}$ seems reasonable if we assume that a two-sided hypothesis about the symbolic distance effect is tested. However, we expect less variance around this mean. Therefore, we decrease the setting for the variation. The wide tails of the Student's *t*-distribution could be one of the reasons for the high range in variation. We change the prior for $\mu_{\gamma}$ to a normal distribution that has more narrow tails. As RTs cannot be negative, we adjust the prior to be positive only by truncating the normal distribution using a lower bound of zero. Finally, we would like to use a more common conjugate prior for the variances to ensure obtaining a proper posterior distribution. ^[Conjugate priors result in posterior distributions that belong to the same distribution family as the prior distribution [@Fink1997]. For a normal likelihood, these are the normal distribution for the mean and the inverse gamma distribution for the variance.] ^[\Revision{However, in rstan, non-conjugate priors are commonly applied} [e.g., @gelman2006prior].] Therefore, we choose an inverse gamma distribution. We can also be more specific about the expected RTs. Based on our knowledge of these types of tasks [e.g., @haaf2017developing], it seems highly plausible that participants will respond within a second or two. We would expect effects in the range of 10 milliseconds to 100 milliseconds. We adjust our priors using these new distribution types and settings, evaluate the prior predictions (shown in panel B of Figure \@ref(fig:priorpredfigurepanel)) and arrive at the following priors:^[Because response times cannot be negative the prior for the overall baseline $\mu_\gamma$ is a truncated normal with a lower bound of zero.]

@@

We follow the same procedure for prior specification in the log-normal model. As before, we prefer to apply conjugate priors. \Revision{The conjugate priors that we applied to the normal model are also conjugate priors for the log-normal model} [@fink1997compendium]. \Revision{For instance, with a normal distribution as prior for the mean of the log-normal model, we assume that the mean of the parameter can be either negative, zero, or positive, and assign a higher probability to values closer to zero. Applying this conjugate prior, the posterior of the mean in the log-normal model will result in a normal distribution as well.} 
As a starting point, we can use the adjusted priors of the normal model and perform a prior prediction, shown in panel C of Figure \@ref(fig:priorpredfigurepanel). These priors result in a slightly higher variance in RTs than we would expect, and, importantly, the mean response time differences are not centered around zero in all conditions. This pattern is inconsistent with our expectations, which means that prior distributions for the log-normal model need to be adjusted.

```{r priorpredfigurepanellog, echo=FALSE, fig.cap="Predicted mean RT difference between digits per repetition. The blue line represents the aggregated trend. Panel A represents the prior prediction based on the priors of the normal model. Panel B shows the prior prediction based on the selected prior for the log-normal model.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="100%", eval = FALSE}
knitr::include_graphics("I - Images/priorpredfigure_lognormal_v2.png")
```
The transformation of the prior distributions is not as simple as the transformation of the parameter itself. Panel C shows that with the same priors as for the normal model, we obtain different data for the log-normal model compared to the normal model (panel B). Therefore, we adjusted the priors until the simulated data matched our expectations and looked like panel B. This yields the following priors that result in the prior prediction shown in panel D of Figure \@ref(fig:priorpredfigurepanel):^[For the overall baseline $\mu_\gamma$ we adjusted the prior based on our expectation of the expected overall response times. A mean parameter of -0.5 for the log-normal likelihood corresponds to a mean response time of 1.]

@@

```{r traceplots3p, echo=FALSE, fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%", fig.cap="Panel A and B show the trace plot of the digit parameter $\\mu_{\\delta_{6}}$ for both packages. Panel C shows the estimated $\\hat{R}$ for all individual effects per parameter group. The stars represent the $\\hat{R}$ of the general effects. Panel D illustrates the variation in the estimated number of effective samples per parameter group. The dashed lines represent the total number of iterations for each package. The stars represent the number of effective samples of the general effects."}
knitr::include_graphics("I - Images/Fig8_traceplots3pnew_v2.png")
```

### Model Fitting 
The model is fitted using the packages `rstan` and `brms`. \Revision{We used different settings for `rstan` and `brms` due to the difference between the models for these packages. Although we tried to make the models as similar as possible, the models are not identical. This is partly due to differences in prior specification (i.e., specifying prior on the variances versus standard deviations).} For `rstan`, four chains are run with 4000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 12000 samples from the joint posterior distribution. For `brms`, four chains are run with 6000 iterations per chain, of which 1000 serve as warm-up. This results in a total of 20000 draws from the joint posterior distribution.

@@

```{r modelposteriorsmufigure3, echo=FALSE, fig.cap="Panel A shows the empirical mean RTs per digit. The standard error is represented by bars around the dots. Panel B and C show the posterior distributions for the general effects estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
knitr::include_graphics("I - Images/Fig9_posteriormu_v3.png")
```

@@

```{r randomparfigure, echo=FALSE, fig.cap="The posterior means for individual effect parameters with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
knitr::include_graphics("I - Images/Fig10_randompar_v2.png")

@@

```{r modelestimatesfigure3, echo=FALSE, fig.cap="Model estimates for digit effect parameters in \\textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates. ", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
knitr::include_graphics("I - Images/Fig11_modelestimatefigure2_onlyrstan_v2.png")
```

@@

```{r traceplotlog, echo=FALSE, fig.cap="(ref:captiontraceplotlog)", fig.align = "left", fig.height = 11, fig.width = 4, out.width="100%"}
knitr::include_graphics("I - Images/Fig12_traceplots3plog_v2.png")
```

### Results
#### General effects  
We inspect the posterior distributions for the general effect parameters presented in Figure \@ref(fig:postplotlog). 
```{r postplotlog, echo=FALSE, fig.cap="The posterior distributions for the general effects in the log-normal model estimated by the two packages. The middle line within the distributions represents the posterior mean. The shaded area within the distributions represent 95\\% credible intervals.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="100%"}
knitr::include_graphics("I - Images/Fig13_posteriormulog_v2.png")
```

@@

```{r randomparplotlog, echo=FALSE, fig.cap="The posterior means for individual effect parameters in the log-normal model with 95\\% credible intervals as estimated by \\textit{rstan}, shown in increasing order. The dashed line represents the general posterior mean. Pink intervals contain zero, blue interval do not contain zero.", fig.align = "left", fig.height = 1, fig.width = 0.5, out.width="100%"}
knitr::include_graphics("I - Images/Fig14_randomparlog_v2.png")
```
We can further inspect the individual estimates for the digit effects in Figure \@ref(fig:modeldigitestplotlog). The figure shows the individual estimates for the digit effects with the distribution of the variability of the individual estimates (not the posterior distribution of the general effect). The figure illustrates that there is considerable variance in the point estimates of the individual digit effects. However, compared to the normal model there is considerably more hierarchical shrinkage. \Revision{This means that in the log-normal model, the individual effects are corrected more towards the group mean, compared to the normal model. Or, the group mean in the log-normal model is not affected as much by individuals showing a divergent trend compared to the normal model. Whether we should prefer a model with more shrinkage depends on the theoretical considerations. If more individual variation is expected, less shrinkage would be preferred. However, the advantage of more shrinkage is that outliers do not influence the estimation of the effect as much. }

@@

```{r modeldigitestplotlog, echo=FALSE, fig.cap="Model estimates for digit effect parameters in the log-normal model in \\textit{rstan}. The points represent the mean parameter estimates for each individual. The violin plots on the right side show the variance of the individual parameter estimates.", fig.align = "left", fig.height = 11, fig.width = 1, out.width="75%"}
knitr::include_graphics("I - Images/Fig15_inddiflogplot_onlyrstan_v2.png")
```

@@

The Bayes factor can also be interpreted as the relative evidence of two competing models. Bayes factors have a few very convenient characteristics. One of them is that Bayes factors are transitive, that is, evidence for the full model over the null model can be obtained by flipping numerator and denominator in equation 13. For a full overview of the advantages of Bayesian inference with Bayes factor, see Wagenmakers et al. [-@wagenmakers2018bayesian].

@@

Imagine this results in $\text{BF}_{\text{nl}} = 100$. This means that the data are 100 times more likely under the normal model than under the log-normal model, clear evidence in favor of the normal model. \Revision{But how do we know whether a Bayes factor of 100-to-1, or 10-to-1 or 3-to-1 fulfills?}. As a reminder, Bayes factors are the relative predictive accuracy of two models for the observed data. Therefore, Bayes factors are odds or ratios. Odds themselves are directly interpretable without the need for decisions or cutoffs. For example, if a presidential candidate is favored 10-to-1 over another one, then these are just the odds, and it is beside the point whether these are large or not. According to this line of thought, Rouder et al. [-@rouder2018theories] argue that researchers should take a similar approach when interpreting and reporting Bayes factors. While there exist rules of thumb to categorize Bayes factors verbally, we suggest researchers use their research context to form substantive conclusions based on Bayes factors, not arbitrary rules of thumb.

@@

```{r sdratioplotall, echo=FALSE, fig.cap="The prior and posterior distribution of the side parameter estimated by the \\textit{brms} package. The dots represent the height of the distributions at zero. The ratio of these dots indicates the Savage-Dickey density ratio. The Bayes factor represents the evidence for the hypothesis that the effect equals zero. A Bayes factor above one indicates that the evidence is in favor of the null hypothesis.", fig.align = "left", fig.height = 6, fig.width = 4, out.width="60%"}
knitr::include_graphics("I - Images/Fig16_sdrplotside_v2.png")
```
With `rstan`, the SDD-ratio can be computed by fitting a density function to the posterior distribution obtained from the MCMC samples, and comparing the value of this density function at $\mu_{\beta} = 0$ to the value of the prior distribution at $\mu_{\beta} = 0$. This is illustrated in the code block below. First, we extract the posterior samples for the parameter $\mu_{\beta}$, the general effect of side. We apply the `logspline` function [@logspline] to the samples to obtain an estimate of the log-density of the posterior distribution. The density at the zero is calculated with the `dlogspline` function. Finally, we compare the estimated density of the posterior to the density of the prior. \textcolor{blue}{This process is explained in further detail by Van Ravenzwaaij and Etz} [-@van2021simulation].

@@

```{r bsbfestimatesbrmsrstan, echo=FALSE, fig.cap="The BF estimates of the other models (0; null, side and digit model) against the full model (1). The green triangles represent the estimates using \\textit{rstan}. The orange dots represent the \\textit{brms} estimates. The dashed line represents a $\\text{BF}_{01}$ of 1. ", fig.align = "left", fig.height = 11, fig.width = 4, out.width="70%"}
knitr::include_graphics("I - Images/Fig17_normalbfplot01.png")
```
Finally, it is time to answer our research question: Is there a symbolic distance effect? When considering all Bayes factor estimates from Figure \@ref(fig:sdratioplotall) and \@ref(fig:bsbfestimatesbrmsrstan), the null model ($\text{BF}_{max} = 232 * 10^{105}$ for `rstan`) performs best compared to the side model ($\text{BF}_{max} = 458 * 10^{68}$ for `rstan`), the digit model ($\text{BF}_{max} = 269 * 10^{31}$ for `rstan`). Therefore, the data from Rouder et al. (2005) show evidence against the symbolic distance effect.

@@

### \Revision{Log-normal Model}
\Revision{We can take a similar approach for the log-normal model to compare the hierarchical models on varying complexity. We evaluate their performance in two scenarios that correspond to scenario 1 and 3 for the normal model. This means that in scenario 1 for the log-normal model, we estimated a null, digit, side and full model based on the prior specified in Equation, specifying large individual differences. For scenario 2, we remove individual variability from the side, digit, and full model. This resulted in the following posterior model probabilities shown in Table} \@ref(tab:postmodproblog).

@@

\Revision{The table shows that the digit model without individual digit effects from scenario 2 has the highest posterior model probability. This is in line with our findings for the normal model. Finally, we can compare the normal model with the highest marginal likelihood to the log-normal with the highest marginal likelihood. The normal digit model from scenario 3 to the log-normal model from scenario 2. The log marginal likelihood for the normal digit model without individual differences was $6441.04$ while the log marginal likelihood for the log-normal digit model without individual differences was $10637.45$. This indicates that the Bayes factor is in favor of the log-normal digit model over the normal digit model.}

The examples illustrate the benefits of comparing hierarchical models of varying complexity. While individual differences are a common phenomenon in psychology, there are instances where general effects are sufficient to describe the data adequately, or where only a random intercept, but no random slopes are necessary. As shown in Tables} \@ref(tab:postmodprob) and \@ref(tab:postmodproblog), Bayesian model comparisons can help researchers to identify the model that best describes the data.

@@

This tutorial showed the application of two R packages: `rstan` and `brms`. Which package to use depends mostly on the complexity of the model and your technical background. For the non-technical user, we recommend using `brms` because of its intuitive model specification. However, the default prior setup can be unsuitable for a specific research context and \Revision{(virtually) always needs} to be adapted. Additionally, users are limited to predefined parametrizations. `rstan` offers the most flexibility in model and prior specification. This flexibility comes with a cost: if the prior or the model is not well defined, results can be misleading.
